---
title: "NHIS_ANALYSIS"
output: html_document
date: '2022-03-21'
Author: 'Kairuo Yan'
---
```{r setup, include=FALSE}
# load libraries of packages
library(tidyverse)
library(ggplot2)
library(dplyr)
library(cluster)
library(factoextra)


```

```{r}
#read files into dataset
health_d <- read.csv(file="adult20.csv")
```


```{r}
#partition a test file for pca exam. Keep the colummn with no more than 10% NA
tested=health_d[, which(colMeans(!is.na(health_d)) > 0.9)]
```

```{r}
#Conduct pca exam and selected features with high evigenvalues
pca = prcomp(na.omit(tested), center = TRUE, scale = FALSE)
#summary(pca)
pcat = as.data.frame(-pca$x[,6:8])
```



```{r}
#etermine the value of k using the elbow method->gap stat chart
fviz_nbclust(pcat, kmeans, method = 'gap_stat')
```


```{r}
#Apply k-means to cluster the data
k = 2
kmeans = kmeans(pcat, centers = k, nstart = 50)
fviz_cluster(kmeans, data = pcat)

```

#Anxiety ANOVA Tests
1. anxiety frequency and age, race and marital stage
```{r}
AX_1 <- aov(ANXFREQ_A ~ AGEP_A+ORIENT_A+RACEALLP_A+SEX_A+MARITAL_A+LEGMSTAT_A, data = health_d)

summary(AX_1)

```

2.anxiety frequency and income and poverty
```{r}
AX_2 <- aov(ANXFREQ_A ~ INCWRKO_A+ INCSSRR_A+EDUC_A+FAMINCTC_A + POVRATTC_A+INCGRP_A+URBRRL+REGION, data = health_d)

summary(AX_2)
```

3.anxiety frequency and smoke habits
```{r}
AX_3 <- aov(ANXFREQ_A ~ SMKAGE_A+CIGNOW_A+ECIGNOW_A+SMKECIGST_A+CIGARCUR_A, data = health_d)

summary(AX_3)

```

4. anxiety frequency and sleep, support and working distance
```{r}
AX_4 <- aov(ANXFREQ_A ~ ASEV_A+CANEV_A+EMPWKHRS2_A+SLPHOURS_A+SLPFLL_A+SUPPORT_A+WRKCLSSD_A+CLSWRKSD_A, data = health_d)

summary(AX_4)


```
5. anxiety frequency and drinking habits
```{r}
AX_5 <- aov(ANXFREQ_A ~ DRKLIFE_A+DRK12MN_A+DRK12MN1_A+DRK12MTP_A+DRK12MTP1_A+DRK12MWK_A+DRK12MYR_A+DRKAVG12M_A+DRKAVG12M1_A+DRKSTAT_A, data = health_d)

summary(AX_5)

```


#Depression ANOVA Tests
1. depression frequency and age, race and marital stage
```{r}
DP_1 <- aov(DEPFREQ_A ~ AGEP_A+ORIENT_A+RACEALLP_A+SEX_A+MARITAL_A+LEGMSTAT_A, data = health_d)

summary(DP_1)

```

2.depression frequency and income and poverty
```{r}
DP_2 <- aov(DEPFREQ_A ~ INCWRKO_A+ INCSSRR_A+EDUC_A+FAMINCTC_A + POVRATTC_A+INCGRP_A+URBRRL+REGION, data = health_d)

summary(DP_2)
```

3.depression frequency and smoke habits
```{r}
AX_3 <- aov(DEPFREQ_A ~ SMKAGE_A+SMKNOW_A+CIGNOW_A+ECIGNOW_A+SMKECIGST_A+CIGARCUR_A, data = health_d)

summary(AX_3)

```

4. depression frequency and sleep, support and working distance
```{r}
AX_4 <- aov(DEPFREQ_A ~ ASEV_A+CANEV_A+EMPWKHRS2_A+SLPHOURS_A+SLPFLL_A+SUPPORT_A+WRKCLSSD_A+CLSWRKSD_A, data = health_d)

summary(AX_4)


```
5. depression frequency and drinking habits
```{r}
AX_5 <- aov(DEPFREQ_A ~ DRKLIFE_A+DRK12MN_A+DRK12MN1_A+DRK12MTP_A+DRK12MTP1_A+DRK12MWK_A+DRK12MYR_A+DRKAVG12M_A+DRKAVG12M1_A+DRKSTAT_A, data = health_d)

summary(AX_5)

```



# Scatterplot for Anxiety vs. Drinking Status
```{r}
he=subset(health_d, health_d$DRKSTAT_A != 10) # delete the answers that are unsure(which is irrelevant to our hypothesis)
h=subset(he, he$DRKSTAT_A != 9) # delete the answers that are unsure(which is irrelevant to our hypothesis)
x <- h$DRKSTAT_A # set x values
y <- h$ANXFREQ_A #set y values



par(font.axis=1, cex.axis=0.6) 
plot(x, y,xaxt="n",yaxt = "n",main = "Anxiety vs. Drinking Status",
     xlab = "Drinking Status", ylab = "Anxiety frequency", col= "olivedrab3", frame = FALSE)       # plot the sctterplot

box(bty = "l")
axis(2, at = 1:5, labels=c("Daily","Weekly","Monthly","A few\ntimes\na year","Never"),las=1)    # label the y-ticks on y-axis

xi=c("Lifetime\nabstainer","Former\ninfrequent","Former\nregular","Former,\n unknown frequency","Current\ninfrequent","Current\nlight", "Current\nmoderate","Current\nheavier","Current\ndrinkerlevel\nunknown","Drinking \nstatus\nunknown")    # label the x-ticks on x-axis
axis(1, at = 1:10, labels=xi,las=1)

abline(lm(y ~ x, data = health_d), col = "orange",lwd=3, lty=2) # add a trend line


```


# Scatterplot for Depression vs. Drinking Status
```{r}
he=subset(health_d, health_d$DRKSTAT_A != 10) # delete the answers that are unsure(which is irrelevant to our hypothesis)
h=subset(he, he$DRKSTAT_A != 9) # delete the answers that are unsure(which is irrelevant to our hypothesis)
x <- h$DRKSTAT_A  #set x values
y <- h$DEPFREQ_A  #set y values



par(font.axis=1, cex.axis=0.6)
plot(x, y,xaxt="n",yaxt = "n",main = "Depression vs. Drinking Status",
     xlab = "Drinking Status", ylab = "Depression frequency", col= "olivedrab3", frame = FALSE)   #plot scatterplot

box(bty = "l")
axis(2, at = 1:5, labels=c("Daily","Weekly","Monthly","A few\ntimes\na year","Never"),las=1)    # label the y-ticks on y-axis

xi=c("Lifetime\nabstainer","Former\ninfrequent","Former\nregular","Former,\n unknown frequency","Current\ninfrequent","Current\nlight", "Current\nmoderate","Current\nheavier","Current\ndrinkerlevel\nunknown","Drinking \nstatus\nunknown")      # label the x-ticks on x-axis
axis(1, at = 1:10, labels=xi,las=1)

abline(lm(y ~ x, data = health_d), col = "orange",lwd=3, lty=2)  # add a trend line


```



#Scatterplot for Anxiety vs. Poverty Ratio
```{r}
#set x, y values
x <- health_d$POVRATTC_A
y <- health_d$ANXFREQ_A


par(font.axis=1, cex.axis=0.6)

#plot scatterplot
plot(x, y,yaxt = "n",main = "Anxiety vs. Poverty Ratio",
     xlab = "Poverty ratio (ratio of the family’s income to the appropriate Federal
poverty threshold)", ylab = "Anxiety frequency", col= "wheat", frame = FALSE)

box(bty = "l")
axis(2, at = 1:5, labels=c("Daily","Weekly","Monthly","A few\ntimes\na year","Never"),las=1)  #set x-ticks for x-axis


# add a trend line
abline(lm(y ~ x, data = health_d), col = "plum3",lwd=3, lty=2)

```


#Scatterplot for Depression vs. Poverty Ratio
```{r}

#set x, y values
x <- health_d$POVRATTC_A
y <- health_d$DEPFREQ_A


par(font.axis=1, cex.axis=0.6)


#plot scatterplot
plot(x, y,yaxt = "n",main = "Depression vs. Poverty Ratio",
     xlab = "Poverty ratio (ratio of the family’s income to the appropriate Federal
poverty threshold)", ylab = "Depression frequency", col= "wheat", frame = FALSE)

box(bty = "l")
axis(2, at = 1:5, labels=c("Daily","Weekly","Monthly","A few\ntimes\na year","Never"),las=1) #set x-ticks for x-axis


# add a trend line
abline(lm(y ~ x, data = health_d), col = "plum3",lwd=3, lty=2)

```





